---
title: "Predictions using the Weight Lifting Exercises Dataset"
author: "Devarshi Sharma"
date: "24/10/2020"
output: html_document
---

Background
----------
Using devices such as _Jawbone Up_, _Nike FuelBand_, and _Fitbit_ it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devicesare part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly improves their health, to find patterns in their behavior, or because they are geeks.  In this project, the goal will be to use data from accelerometers on the belts of 6 participants. They performed barrell lifts correctly and incorrectly 5 different ways

The goal of this project is to predict the manner in which they performed the exercise.

<hr />

Data Set
--------
The training data for this project are available here:
[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv)

The test data are available here:
[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)

More information is available from the website here: [http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har](http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har) (see the section on the Weight Lifting Exercise Dataset).
If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

<hr />

Submission
-----------

#### Loading Data set
```{r}
if(!file.exists("pml-training.csv"))
{
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "pml-training.csv", method = 'curl')
}
dataset <- read.csv("pml-training.csv", na.strings = c("NA", ""))
if(!file.exists("pml-testing.csv"))
{
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", "pml-testing.csv", method = 'curl')
}
val <- read.csv("pml-testing.csv")
```

### Importing Librarie 
```{r warning=FALSE, error=FALSE}
library(caret)
library(randomForest)
```

### Setting Seed, so that analysis is reproducible
```{r warning=FALSE, error=FALSE}
set.seed(100)
```

### Dividing the data set into 60/40 training and testing set
```{r warning=FALSE, error=FALSE}
inTrain = createDataPartition(y=dataset$classe, p=0.6, list=FALSE)
train = dataset[inTrain,]
test = dataset[-inTrain,]
```

### Removing all Null values 
```{r warning=FALSE, error=FALSE}
NullCols = sapply(train, function(x) {sum(is.na(x))}) 
NullCols
colsWithNull = names(NullCols[NullCols > 0])
train = train[, !names(train) %in% colsWithNull] 
names(train)
train <- train[, !names(train) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
```

### Creating a Validation set 
```{r warning=FALSE, error=FALSE}
NullCols = sapply(val, function(x) {sum(is.na(x))})
colsWithNull = names(NullCols[NullCols > 0])
val= val[, !names(val) %in% colsWithNull]
val <- val[, !names(val) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
```

### Creating testing set 
```{r warning=FALSE, error=FALSE}
NullCols = sapply(test, function(x) {sum(is.na(x))})
colsWithNull = names(NullCols[NullCols > 0])
test = test[, !names(test) %in% colsWithNull]
test <- test[, !names(test) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
```

### Creating model by using Decision tree 
```{r warning=FALSE, error=FALSE}
modFitDT <- rpart(classe ~ ., data = train, method = "class")
predictDT <- predict(modFitDT, test, type = "class")
rpart.plot(modFitDT, main="Classification Tree", extra=102, under=TRUE, faclen=0, cex = 0.55)
confusionMatrix(factor(predictDT), factor(test$classe))
```

### Creating model by using RandomForest method
```{r warning=FALSE, error=FALSE}
model <- randomForest(as.factor(classe) ~ .,   data=train, ntree = 100)
pred <- predict(model, test)
confusionMatrix(factor(pred), factor(test$classe))
modelAcc <- confusionMatrix(factor(pred), factor(test$classe))$overall[[1]]
```

<hr />

Conclusion
----------

We can conclude by saying that RandomForest gives more accurate results than Decision Tree. Hence, for this problem we use RandomForest to predict the manner in which the participants did the exercise. 


We conclude that Random forrest gives more accurate results than Decision tree. Thus, for this probkem we should use random forrest to predict the exercise the participants did 
<hr />
<br>